### 2025-02-09 analysis/analysis.qmd
feat(analysis.qmd): update classification analyses

## Classification studies

@@

m.cor <- metacor(cor = valuesMax,     # values 
                 n = stimulus_n,
                 studlab = studyREF, # unique_id
                 data = C_summary,
                 common = FALSE,
                 random = TRUE,
                 prediction = TRUE,
                 backtransf = TRUE,
                 sm = "ZCOR",
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Classification: Summary")

#m.cor_backtransformed <- m.cor
#m.cor_backtransformed$TE <- FisherZInv(m.cor_backtransformed$TE)
print(funnel(m.cor, common = FALSE, studlab=TRUE,backtransf=TRUE))

```
#### Custom plot

```{r}
library(ggrepel)
tmpdata <- data.frame(SE = FisherZInv(m.cor$seTE), Zr = FisherZInv(m.cor$TE),studies=m.cor$studlab)

@@

@@

```{r}
forest(m.cor,
       sortvar = TE,
       prediction = FALSE, 
             print.tau2 = FALSE,
             leftlabs = c("Author", "N"),studlab = citekey)

```
#### By class n = 2 vs. > 2
```{r}
C_summary$classes <- 0
C_summary[C_summary$n_classes < 3,]$classes <- "2"
C_summary[C_summary$n_classes >= 3,]$classes <- ">2"

m.cor <- metacor(cor = valuesMax,     # values 
                 n = stimulus_n, # could also change to n_classes
                 studlab = studyREF, # unique_id
                 data = C_summary,
                 subgroup = classes,
                 common = FALSE,
                 random = TRUE,
                 prediction = TRUE,
                 backtransf = TRUE,
                 sm = "ZCOR",
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Classification: Summary")

forest(m.cor,
       sortvar = TE,
       prediction = FALSE, 
             print.tau2 = FALSE,
             leftlabs = c("Author", "N"),studlab = citekey)

@@

#### Explore qualities (in progress)

```{r}
O <- find.outliers(m.cor)
# 6 datasets identified as outliers, without them the r drops to 0.5781
#infan <- InfluenceAnalysis(m.cor)
#print(eggers.test(m.cor))

#### Re-run the analysis without the outliers

```{r}
#| eval: false
#| echo: false
outliers <- c("")

tmp2<-tmp[!tmp$unique_id %in% outliers,]
m.cor <- metacor(cor = values, 
                 n = stimulus_n,
                 studlab = unique_id, # unique_id
                 data = tmp2,
                 common = FALSE,
                 random = TRUE,
                 sm = "ZCOR",
                 prediction = TRUE,
                 backtransf = TRUE,
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Regression: Valence: Summary")

fp <- funnel(m.cor, common = TRUE,studlab=FALSE)

@@

```{r}
#| eval: true
tmpdata <- data.frame(SE = m.cor$seTE, Zr = m.cor$TE, studies=m.cor$studlab)

@@

estimate = m.cor$TE.common
se = m.cor$seTE.common
se.seq=seq(0, max(m.cor$cor), 0.001)
ll95 = estimate-(1.96*se.seq)
ul95 = estimate+(1.96*se.seq)
ll99 = estimate-(3.29*se.seq)
ul99 = estimate+(3.29*se.seq)
meanll95 = estimate-(1.96*se)
meanul95 = estimate+(1.96*se)
dfCI = data.frame(ll95, ul95, ll99, ul99, se.seq, estimate, meanll95, meanul95)

fp = ggplot(NULL) +
  geom_point(aes(x = SE, y = Zr), color='grey50',data=tmpdata) +
  geom_text_repel(aes(x = SE, y = Zr, label=studyREF), data=tmpdata,size=2.5,max.overlaps = 40)+
  xlab('Standard Error') + ylab('Fisher\'s z transformed correlation')+
  geom_line(aes(x = se.seq, y = ll95), linetype = 'dotted', data = dfCI) +
  geom_line(aes(x = se.seq, y = ul95), linetype = 'dotted', data = dfCI) +
  geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI) +
  geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI) +
#  scale_x_reverse()+
  scale_x_reverse(breaks=seq(0,0.2,0.05),limits=c(0.15,0),expand=c(0.00,0.00))+
  scale_y_continuous(breaks=seq(0.3,1.25,0.20),limits=c(0.3,1.23))+
  coord_flip()+
  theme_bw()
print(fp)

@@

#### Subgroup analysis according to techniques

add `journal_type` and `stimulus_genre_mixed` as a grouping option

```{r}
#| eval: false
m.cor_subgroups <- update(m.cor, 
       subgroup = model_class_id, 
#       subgroup = journal_type, 
#       subgroup = stimulus_genre_mixed, 
       tau.common = FALSE,
       prediction = TRUE)

>>>>

## Classification studies: overall success

@@

m.cor.c.all <- metacor(cor = values,     # values 
                 n = stimulus_n,
                 studlab = studyREF, # unique_id
                 data = C_studies,
                 common = FALSE,
                 random = TRUE,
                 prediction = TRUE,
                 backtransf = TRUE,
                 sm = "ZCOR",
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Classification: Summary")

m.cor.c.all

#m.cor_backtransformed <- m.cor
#m.cor_backtransformed$TE <- FisherZInv(m.cor_backtransformed$TE)
print(funnel(m.cor.c.all, common = FALSE, studlab=TRUE,backtransf=TRUE))

```
## Define splits for genre & stimulus n combinations

```{r}
# get stimulus and genre metadata for classification studies:
C_studies |>
  group_by(citekey) |>
  summarize(feature_n = unique(feature_n),
            feature_n_complexity = unique(feature_n_complexity),
            stimulus_genre_mixed = unique(stimulus_genre_mixed)) -> C_splits

# create feature_n_complexity_genre column:
C_splits$feature_n_complexity_genre <- ""

# define splits
C_splits[C_splits$feature_n_complexity %in% "Feature n > 30 & < 300" &
        C_splits$stimulus_genre_mixed == "MultiGenre",]$feature_n_complexity_genre <- "Medium multi-genre study"
C_splits[C_splits$stimulus_genre_mixed %in% "SingleGenre" &
                             C_splits$feature_n_complexity %in% "Feature n < 30" ,]$feature_n_complexity_genre <- "Small single-genre study"
C_splits[C_splits$feature_n_complexity == "Feature n > 300",]$feature_n_complexity_genre <- "Huge single or multigenre study"
C_splits[C_splits$citekey == "alvarez2023ri",]$feature_n_complexity_genre <- "Small, multi-genre study"
C_splits[C_splits$citekey == "zhang2016br",]$feature_n_complexity_genre <- "Medium, single-genre study"
# add splits to summary:
C_summary <- left_join(C_summary, C_splits)

### Define splits for binary vs. multiclass

```{r}
C_summary$classes <- 0
C_summary[C_summary$n_classes < 3,]$classes <- "Binary"
C_summary[C_summary$n_classes >= 3,]$classes <- "Multiclass"

## Classification models: best models

```{r}
m.cor.c <- metacor(cor = valuesMax,     # values 
                 n = stimulus_n,
                 studlab = studyREF, # unique_id
                 data = C_summary,
                 common = FALSE,
                 random = TRUE,
                 prediction = TRUE,
                 backtransf = TRUE,
                 sm = "ZCOR",
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Classification: Summary")

### Custom funnel for classification studies

```{r}
#| fig-width: 12
#| fig-height: 6
fig3a <- forest(m.cor.c,
                sortvar = TE,
                prediction = FALSE, 
                print.tau2 = FALSE,
                leftlabs = c("Author", "N"),studlab = studyREF)

fig3b<-funnel(m.cor.c, common = FALSE, studlab=TRUE,backtransf=TRUE)

library(forestplot)
data<-tibble::tibble(mean=m.cor.c$cor,lower=FisherZInv(m.cor.c$lower),upper=FisherZInv(m.cor.c$upper),study=m.cor.c$studlab,n=m.cor.c$n,cor=round(m.cor.c$cor,2))
data<-dplyr::arrange(data,mean)

fp3 <- grid.grabExpr(print(data |>
                             forestplot(labeltext = c(study, n, cor),
                                        xlab = "Correlation",
                                        xticks = c(0, .25,.5,.75, 1),
                                        clip = c(0, 1))|>
                             fp_add_header(study = "Study",n = "N",cor = expression(italic(r))) |>
                             fp_append_row(mean  = m.cor.c$TE.common,
                                           lower = m.cor.c$lower.common,
                                           upper = m.cor.c$upper.common,
                                           study = "Summary",
                                           n = sum(m.cor.c$n),
                                           cor = round(m.cor.c$TE.common,2),
                                           is.summary = TRUE) |>
                             fp_set_style(box = "grey50",
                                          line = "grey20",
                                          summary = "black",
                                          txt_gp = fpTxtGp(label = list(gpar(cex = 0.80)),
                                                           ticks = gpar(cex = 0.80),
                                                           xlab  = gpar(cex = 0.80)))|>
                             fp_decorate_graph(grid = structure( m.cor.c$TE.common,gp = gpar(lty = 2, col = "grey30")))
)
)

source('../etc/custom_funnel_plot.R')
fp4 <- custom_funnel_plot(m.cor.c)

gridExtra::grid.arrange(fp3, fp4, ncol=2, widths=c(2,1),heights=c(2,1))

#### Custom plot

```{r}
library(ggrepel)
tmpdata <- data.frame(SE = FisherZInv(m.cor.c$seTE), Zr = FisherZInv(m.cor.c$TE),studies=m.cor.c$studlab)

@@

## Exploring heterogeneity

```{r}
#| eval: true
#| echo: true
# Method 1: Give us the Egger's test about beta coefficient from funnel
print(eggers.test(m.cor.c))

# Method 2: Find the impact to the results when removing those outside the 95CI
O <- find.outliers(m.cor.c) # 13 remaining out of 24
print(O)
# Method 3: Leave-out-out analysis etc for individual influence (not useful here)
#infan <- InfluenceAnalysis(m.cor)

# Method 4: Focus on 10% of most precise studies (Stanley, Jarrel, Doucouliagos 2010)
thres<-quantile(m.cor.c$seTE,0.1)
ind<-m.cor.c$seTE<=as.numeric(thres)
m.cor10pct <- update(m.cor.c, subset = which(ind))

# Method 5: P curve analysis (Simonsohn, Simmons & Nelson, 2015)
pcurve(m.cor.c)
```
### Re-run the analysis without outliers

```{r}
#| eval: true
#| echo: false
outliers <- O$out.study.random

tmp2<-C_summary[!C_summary$studyREF %in% outliers,]
m.cor.c.o <- metacor(cor = valuesMax, 
                 n = stimulus_n,
                 studlab = tmp2$studyREF,
                 data = tmp2,
                 common = FALSE,
                 random = TRUE,
                 sm = "ZCOR",
                 prediction = TRUE,
                 backtransf = TRUE,
                 method.tau = "REML",# could be PM (Paule-Mandel) as well
                 method.random.ci = "HK", 
                 title = "MER: Regression: Valence: Summary")

@@

```{r}
forest(m.cor.c.o,
       sortvar = TE,
       prediction = FALSE, 
             print.tau2 = FALSE,
             leftlabs = c("Author", "N"),studlab = citekey)

funnel(m.cor.c.o, common = FALSE, studlab=TRUE,backtransf=TRUE)

```
## Subgroup analyses

### By binary vs. multi-class classification

```{r}
m.cor_classes <- update(m.cor.c, 
       subgroup = C_summary$classes, 
       tau.common = FALSE,
       prediction = TRUE)

forest(m.cor_classes,
       sortvar = TE,
       prediction = FALSE, 
             print.tau2 = FALSE,
             leftlabs = c("Author", "N"),studlab = citekey)

@@

### By single vs. multi-genre

```{r}
#| eval: false
m.cor_subgroups <- update(m.cor.c, 
       subgroup = stimulus_genre_mixed, 
       tau.common = FALSE,
       prediction = TRUE)

### By genre + feature complexity

```{r}
m.cor_complexity <- update(m.cor.c, 
       subgroup = C_summary$feature_n_complexity_genre, 
       tau.common = FALSE,
       prediction = TRUE)

print(m.cor_complexity)
```
### By model type

```{r}
#| eval: false
m.cor_subgroups <- update(m.cor.c, 
       subgroup = model_class_id, 
       tau.common = FALSE,
       prediction = TRUE)

@@

```{r}
#| eval: true
tmpdata <- data.frame(SE = m.cor.c$seTE, Zr = m.cor.c$TE, studies=m.cor.c$studlab)

@@

estimate = m.cor.c$TE.common
se = m.cor.c$seTE.common
se.seq=seq(0, max(m.cor.c$cor), 0.001)
ll95 = estimate-(1.96*se.seq)
ul95 = estimate+(1.96*se.seq)
ll99 = estimate-(3.29*se.seq)
ul99 = estimate+(3.29*se.seq)
meanll95 = estimate-(1.96*se)
meanul95 = estimate+(1.96*se)
dfCI = data.frame(ll95, ul95, ll99, ul99, se.seq, estimate, meanll95, meanul95)

fp = ggplot(NULL) +
  geom_point(aes(x = SE, y = Zr), color='grey50',data=tmpdata) +
  geom_text_repel(aes(x = SE, y = Zr, label=studyREF), data=tmpdata,size=2.5,max.overlaps = 40)+
  xlab('Standard Error') + ylab('Fisher\'s z transformed correlation')+
  geom_line(aes(x = se.seq, y = ll95), linetype = 'dotted', data = dfCI) +
  geom_line(aes(x = se.seq, y = ul95), linetype = 'dotted', data = dfCI) +
  geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI) +
  geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI) +
#  scale_x_reverse()+
  scale_x_reverse(breaks=seq(0,0.2,0.05),limits=c(0.15,0),expand=c(0.00,0.00))+
 # scale_y_continuous(breaks=seq(0.3,1.25,0.20),limits=c(0.3,1.23))+
  coord_flip()+
  theme_bw()
print(fp)

@@
